Inference methods comparison for the simplified Naomi model using Pareto-smoothed importance sampling

### 2023-05-10 src/naomi-simple_psis/psis.Rmd
Get NLL for all methods, step of PSIS #41

```{r}
tmb <- readRDS("depends/tmb.rds")
aghq <- readRDS("depends/aghq.rds")
tmbstan <- readRDS("depends/tmbstan.rds")

depends <- yaml::read_yaml("orderly.yml")$depends
```

@@

```{r}
stopifnot(names(tmb$fit$sample) == names(aghq$quad$sample))
stopifnot(names(tmb$fit$sample) == names(adam$adam$sample))
stopifnot(names(tmb$fit$sample) == names(tmbstan$mcmc$sample))
```

@@

Although we can extract a `TMB` objective function from `tmb` directly, it would correspond to the Laplace approximation.
As such we use a new objective function using `naomi_simple.cpp` directly:

```{r}
objfull <- readRDS("depends/objfull.rds")
objfull$fn(objfull$par)
```

@@

```{r}
tmbstan_samples <- rstan::extract(tmbstan$mcmc$stanfit)

data.frame(
  y = tmbstan_samples$lp__,
  x = 1:length(tmbstan_samples$lp__)
) %>%
  ggplot(aes(x = x, y = y)) +
    geom_point(alpha = 0.4) +
    labs(y = "Log-likelihood", x = "Draw number") +
    theme_minimal() 
```

@@

```{r}
objfull$par

#' TODO
```

>>>>

```{r}
tmb <- readRDS("depends/tmb.rds")
aghq <- readRDS("depends/aghq.rds")
tmbstan <- readRDS("depends/tmbstan.rds")å
```

@@

```{r}
stopifnot(names(tmb$fit$sample) == names(aghq$quad$sample))
stopifnot(names(tmb$fit$sample) == names(tmbstan$mcmc$sample))
```

@@

Although we can extract a `TMB` objective function from `tmb` directly, it would correspond to the Laplace approximation.
Instead we use `objfull` obtained from a previous report, and load the `naomi_simple` DLL:

```{r}
TMB::compile("naomi_simple.cpp")
dyn.load(TMB::dynlib("naomi_simple"))

objfull <- readRDS("depends/objfull.rds")
objfull$fn(objfull$par)
```

@@

```{r}
tmbstan_extract <- rstan::extract(tmbstan$mcmc$stanfit)

data.frame(
  y = tmbstan_extract$lp__,
  x = 1:length(tmbstan_extract$lp__)
) %>%
  ggplot(aes(x = x, y = y)) +
    geom_point(alpha = 0.4) +
    labs(y = "Log-likelihood", x = "Draw number") +
    theme_minimal()
```

@@

```{r}
par_samp_matrix <- function(sample) {
  # Remove the Naomi outputs, ending with either _out or _ll
  x <- sample[!(stringr::str_ends(names(sample), "_out") | stringr::str_ends(names(sample), "_ll"))]
  # Reorder to be the same as objfull$par
  x <- x[unique(names(objfull$par))]
  # Bind rows together
  do.call(rbind, x)
}

tmb_samples <- par_samp_matrix(tmb$fit$sample)
aghq_samples <- par_samp_matrix(aghq$quad$sample)
tmbstan_samples <- par_samp_matrix(tmbstan$mcmc$sample)

tmb_ll <- apply(tmb_samples, 2, FUN = function(x) -1 * objfull$fn(x))
aghq_ll <- apply(aghq_samples, 2, FUN = function(x) -1 * objfull$fn(x))
tmbstan_ll <- apply(tmbstan_samples, 2, FUN = function(x) -1 * objfull$fn(x))

data.frame(
  y = c(tmb_ll, aghq_ll, tmbstan_ll),
  x = c(1:length(tmb_ll), 1:length(aghq_ll), 1:length(tmbstan_ll)),
  method = c(rep("TMB", times = length(tmb_ll)), rep("aghq", times = length(aghq_ll)), rep("tmbstan", times = length(tmbstan_ll)))
) %>%
  ggplot(aes(x = x, y = y, col = method)) +
    geom_point(alpha = 0.4) +
    facet_grid(~ method, scales = "free_x") +
    scale_colour_manual(values = c("#56B4E9", "#009E73", "#E69F00")) +
    labs(y = "Log-likelihood", x = "Draw number", col = "Method") +
    theme_minimal()

#' Check that these are the same
plot(sort(tmbstan_extract$lp__), sort(tmbstan_ll))

-1 * mean(tmb_ll)
-1 * mean(aghq_ll)
-1 * mean(tmbstan_ll)

#' Run PSIS
#' Need to get log-likelihood under the proposal
#' A Gaussian for TMB
#' A mixture of Gaussians for aghq
lw <- tmb_ll
loo::psis(log_ratios = lw, r_eff = 1)

lw <- aghq_ll
loo::psis(log_ratios = lw, r_eff = 1)
```
